A Novel Model-Free Actor-Critic Reinforcement Learning Approach for Dynamic Target Tracking

### 2020-07-10 paperMIC2020/paperIEEE-MIC2020-V1.tex
addressed issues

Fig.~\ref{fig:leaderFollowerSetup} shows the setup of the problem that we address in this manuscript. The robot is characterized by its 2D position $(x_k,y_k)$ and orientation $\theta_k\in[-\pi,\pi)$ with respect to the global coordinate frame X-Y at discrete time index $k=0,1,\ldots,$ where time $t=kT$ with $T>0$ being the sampling time. The time-varying position of the target is denoted by the position vector ${\bf p}_k^{[d]}= [x_k^{[d]},y_k^{[d]}]^T\in\mathbb{R}^2.$  Let the position and orientation (pose) of the robot is denoted by the vector  ${\bf q}_k^T\equiv[x_k,y_k,\theta_k]\in\mathbb{R}^2\times\mathbb{S}^1.$ Without loss of generality, assume that the robot tracks the target trajectory while maintaining line-of-sight distance (also known as the safe distance between the robot and the target) $d>0.$ We emphasize that the mathematical models of both robot and the target are not required in the proposed control technology. For illustration, let us assume that the robot (rear-wheel drive) follows the discrete-time kinematic model described by %
%
\begin{subequations}
  \label{eq:robotModelDT}  
  \begin{align}
    x_{k+1} &= x_k +T\nu_k\cos\theta_k,\\
    y_{k+1} &= y_k +T\nu_k\sin\theta_k,\\
    \theta_{k+1} &= \theta_k + T\frac{\nu_k}{L}\tan\gamma_k,
  \end{align}
\end{subequations}
%
where $\gamma_k\in(-\frac{\pi}{2},\frac{\pi}{2})$ is the steering angle of the font wheel/castor of the robot (see Fig.~\ref{fig:leaderFollowerSetup}), and $\nu_k\in\mathbb{R}$ is the linear speed. Without loss of generality, this work considers a target modeled by an integrator. The discrete-time kinematic model of the target is given as %
%
 \begin{align}
   \label{eq:leaderDT}
   \mathbf{p}_{k+1}^{[d]} = \mathbf{p}_k^{[d]} + T_s \, \mathbf{u}_k^{[d]},
 \end{align}
% 
 with $\mathbf{u}_k^{[d]}\in\mathbb{R}^2$ being the control input vector (velocity) that defines the trajectory (considered random in this work) of the target.  In case, the robot modeled by~\eqref{eq:robotModelDT} is described by a differential drive mobile robot, then its left-wheel velocity $\nu_{L,k}$ and right-wheel velocity $\nu_{R,k}$ are related by
 %
  \begin{figure}
   \centering
   \fcolorbox{gray!10}{gray!5}{
     \includegraphics[width=0.45\textwidth]{figs/ipe/LFMICSetup.eps}
    }
   \caption{Mobile robot and its dynamic target tracking problem setup. \todo[inline]{In this figure, replace $\tilde\theta_k$ with $\theta_k^{'}$}}
   \label{fig:leaderFollowerSetup}
 \end{figure}
 %
 \begin{subequations}
   \begin{align}
     \nu_{k} &= \frac{1}{2}(\nu_{R,k} + \nu_{L,k}),\\
     \omega_{k} &= \frac{1}{\ell}(\nu_{R,k} - \nu_{L,k}) = \frac{\nu_{k}}{L}\tan(\gamma_k) 	
   \end{align}
 \label{eq:robotModel1-DT}
 \end{subequations}
 %
 where $\omega_k = \frac{\nu_k}{L}\tan\gamma_k$ is the angular velocity of the robot's center $(x_k,y_k).$ Let us define the error vector as 
 \begin{align}
     \label{eq:stateError}
   \mathbf{e}_k = \left[ \tilde{\rho_k},\tilde\theta_k \right]^T=\left[ \sqrt{\tilde{x}_k^2+\tilde{y}_k^2},\tilde{\theta_k} \right]^T% = \\
   % \begin{bmatrix}
   %   \sqrt{(x_k^{[d]} - x_k - d\cos\theta_k^{'})^2+
   %   (y_k^{[d]} - y_k - d\sin\theta_k^{'})^2},
   %   \theta_k^{'} - \theta_k
   % \end{bmatrix},
 \end{align}
% 
 where $\theta_k^{'} = \mathrm{atan2}\left(y_k^{[d]}-y_k, x_k^{[d]}-x_k\right)$
 and $\tilde{\rho}_k =\sqrt{\tilde{x}_k^2+\tilde{y}_k^2} $ is the Euclidean distance (position error) at time instant $k$ as shown in Fig.~\ref{fig:leaderFollowerSetup} with $\tilde{x}_k = x_k^{[d]} - x_k - d\cos\theta_k^{'},$ $\tilde{y}_k = y_k^{[d]} - y_k - d\sin\theta_k^{'},$ and $\tilde{\theta}_k = \theta_k^{'} - \theta_k$ is the orientation error.  The control problem can then be formally stated as follows: Find $\nu_k$ and $\gamma_k$ such that ${\bf e}_k\to {\bf 0}$ as  $k\to\infty$ subject to~\eqref{eq:robotModelDT}~and~\eqref{eq:leaderDT}.

@@

@@

 This section adopts the theoretical results discussed in the previous section by simulating Algorithm~\ref{alg:ModelFreeTracking} using a differential drive mobile robot. The proposed actor-critic RL algorithm is tested  utilizing MATLAB software and Robot Operating System (ROS)  as a preliminary step to implement the Algorithm experimentally in real world. Here we present the performance of the proposed method and the convergence characteristics of the actor and critic weights. The weighting matrices are set to $\mathbf{Q} = \mathrm{diag}[0.001,0.001]$ and $\mathbf{R} = \mathrm{diag}[10^{-5}, 10^{-5}].$ % 
  % \[Q=0.001 
  % \begin{bmatrix}
  % 1       & 0   \\
  % 0       & 1   \\
  % \end{bmatrix},
  % R=0.00001
  % \begin{bmatrix}
  % 1       & 0 \\
  % 0       & 1 
  % \end{bmatrix}.\]

@@

 In the first scenario, the differential drive robot was initially placed at position $(x^d,y^d) = (-1,-5)~[\si{\meter}] $ with an orientation of $\theta = 0\degree.$ The target which is modeled as a \comment{cylinder object}{we are running MATLAB simulation and using point robot as a target} was initially located at a random position around  $(x,y) = (0,0)~[\si{\meter}].$ The simulation was run for the duration of $t_f=60~[\second].$ The target is set to move following a sinusoidal path modeled as  $
 x^{[d]}(t) = \alpha t, %
 y^{[d]}(t) = 5 \,\sin(\alpha t),$  and %
 $\theta^{[d]}(t) = cos(\alpha t),
 $ with $\alpha = 10\pi/t_f.$ Simulation results are shown in Fig.~\ref{fig:performanceSin}. In the beginning of the simulation, the robot starts to move away from the target according to the initial random actor weights. As the robot collects more data, the robot adjust the control actions as the actor and critic weights started to converge and the robot asymptotically follows the target's trajectory. %
%
 \begin{figure*}[htbp]%

 \subfigure[][]{%
    \label{fig:trajectorySin}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/trajectoryRandomPathDistance.eps} }
	\subfigure[][]{%    
    \label{fig:euclideanDistanceSin}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/euclideanDistance.eps} }
  \\
\subfigure[][]{%
    \label{fig:stateErrorSin}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/stateErrorRandomPathDistance.eps} }
 \subfigure[][]{%
    \label{fig:controlInputsSin}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/controlInputRandomPathDistance.eps} }
  \\
 \subfigure[][]{%
    \label{fig:criticWeightSin}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/weightRandomPathDistance.eps} }
     \subfigure[][]{%
    \label{fig:actorWeightSin}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/weightActor.eps} }
\caption[Robot tracking a target that is following a sinusoidal trajectory.]{\comment{Robot tracking a target that is  following a sinusoidal trajectory:}{Are you using CoppeliaSim results?} 
  \subref{fig:trajectorySin} robot and target trajectories,~\subref{fig:stateErrorSin} position error $\tilde\rho_k$,~\subref{fig:controlInputsSin} linear speed and steering angle of the target, \subref{fig:criticWeightSin} critic learning weights~$\bm{\omega}_c,$ and \subref{fig:actorWeightSin} actor learning weights~${\bf W}_a$}%
  \label{fig:performanceSin}%
\end{figure*} 
%

In the second scenario, the target is set to move on a completely random trajectory starting from a random initial position $(x^{[d]},y^{[d]})\approx (1,5)~[\si{\meter}],$ where the \comment{target is now modeled as an unicycle robot}{Previously we talk about point robot model for the target but now you are using the following model for the target} described by %
%
 $x_{k+1}^{[d]} = x_k^{[d]} + T\nu_k^{[d]}\cos\theta_k^{[d]},$ %
 $y_{k+1}^{[d]} = y_k^{[d]} + T\nu_k^{[d]}\sin\theta_k^{[d]},$ %
 $\theta_{k+1}^{[d]} = \theta_k^{[d]} +T\omega_k^{[d]},$ %
 %
 where $\nu^{[d]}(t)=0.2~[\si{\meter}]$\comment{~}{What about $\omega$?} The robot initial position and orientation being $(x,y) = (0,0)~[\si{\meter}]$ and $\theta_0 = $ We intensify that the robot has no previous knowledge of the dynamic model of the moving target\comment{nor it's initial position}{Robot does need to know the target's position through the sensors}. The results obtained by the simulation are shown in Fig.~\ref{fig:performanceRandom}. The control actions of the robot are determined by the random actor weights. We clearly see tracking error converges and approaches zero as the control signal converging to values that allows target tracking. The results reveal the ability of the proposed algorithm to successfully track the dynamic moving target and adapt to the changes in the target's trajectory and learns online by trial and error in different scenarios. %
%    
 \begin{figure*}[htbp]%
 \subfigure[][]{%
    \label{fig:trajectoryRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/trajectoryRandomPathDistance.eps} }
	\subfigure[][]{%    
    \label{fig:euclideanDistanceRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/euclideanDistance.eps} }
\subfigure[][]{%
    \label{fig:stateErrorRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/stateErrorRandomPathDistance.eps} }
 \subfigure[][]{%
    \label{fig:controlInputsRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/controlInputRandomPathDistance.eps} }
 \subfigure[][]{%
    \label{fig:criticWeightRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/weightRandomPathDistance.eps} }
     \subfigure[][]{%
    \label{fig:actorWeightRandom}%
    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/weightActor.eps} }
  \caption[Robot tracking a target that is  following a random trajectory.]{Robot tracking a target that is  following a random trajectory:
  \subref{fig:trajectoryRandom} robot and target trajectories,~\subref{fig:stateErrorRandom} position error $\tilde\rho_k$,~\subref{fig:controlInputsRandom} linear speed and steering angle of the target, \subref{fig:criticWeightRandom} critic learning weights~$\bm{\omega}_c,$ and \subref{fig:actorWeightRandom} actor learning weights~${\bf W}_a$}%
  \label{fig:performanceRandom}%
\end{figure*}

>>>>

Fig.~\ref{fig:leaderFollowerSetup} shows the setup of the problem that we address in this manuscript. The robot is characterized by its 2D position $(x_k,y_k)$ and orientation $\theta_k\in[-\pi,\pi)$ with respect to the global coordinate frame X-Y at discrete time index $k=0,1,\ldots,$ where time $t=kT$ with $T>0$ being the sampling time. The time-varying position of the target is denoted by the position vector ${\bf p}_k^{[d]}= [x_k^{[d]},y_k^{[d]}]^T\in\mathbb{R}^2.$  Let the position and orientation (pose) of the robot is denoted by the vector  ${\bf q}_k^T\equiv[x_k,y_k,\theta_k]\in\mathbb{R}^2\times\mathbb{S}^1.$ Without loss of generality, assume that the robot tracks the target trajectory while maintaining line-of-sight distance (also known as the safe distance between the robot and the target) $d>0.$ We emphasize that the mathematical models of both robot and the target are not required in the proposed control technology. For illustration, let us assume that the robot (rear-wheel drive) follows the discrete-time kinematic model described by %
%
\begin{subequations}
  \label{eq:robotModelDT}  
  \begin{align}
    x_{k+1} &= x_k +T\nu_k\cos\theta_k,\\
    y_{k+1} &= y_k +T\nu_k\sin\theta_k,\\
    \theta_{k+1} &= \theta_k + T\frac{\nu_k}{L}\tan\gamma_k,
  \end{align}
\end{subequations}
%
where $\gamma_k\in(-\frac{\pi}{2},\frac{\pi}{2})$ is the steering angle of the font wheel/castor of the robot (see Fig.~\ref{fig:leaderFollowerSetup}), and $\nu_k\in\mathbb{R}$ is the linear speed. Without loss of generality, this work considers a target modeled by an integrator. The discrete-time kinematic model of the target is given as %
%
 \begin{align}
   \label{eq:leaderDT}
   \mathbf{p}_{k+1}^{[d]} = \mathbf{p}_k^{[d]} + T_s \, \mathbf{u}_k^{[d]},
 \end{align}
% 
 with $\mathbf{u}_k^{[d]}\in\mathbb{R}^2$ being the control input vector (velocity) that defines the trajectory (considered random in this work) of the target.  In case, the robot modeled by~\eqref{eq:robotModelDT} is described by a differential drive mobile robot, then its left-wheel velocity $\nu_{L,k}$ and right-wheel velocity $\nu_{R,k}$ are related by
 %
  \begin{figure}
   \centering
   \fcolorbox{gray!10}{gray!5}{
     \includegraphics[width=0.45\textwidth]{figs/ipe/LFMICSetup.eps}
    }
   \caption{Mobile robot and its dynamic target tracking problem setup.}
   \label{fig:leaderFollowerSetup}
 \end{figure}
 %
 \begin{subequations}
   \begin{align}
     \nu_{k} &= \frac{1}{2}(\nu_{R,k} + \nu_{L,k}),\\
     \omega_{k} &= \frac{1}{\ell}(\nu_{R,k} - \nu_{L,k}) = \frac{\nu_{k}}{L}\tan(\gamma_k) 	
   \end{align}
 \label{eq:robotModel1-DT}
 \end{subequations}
 %
 where $\omega_k = \frac{\nu_k}{L}\tan\gamma_k$ is the angular velocity of the robot's center $(x_k,y_k).$ Let us define the error vector as 
 \begin{align}
     \label{eq:stateError}
   \mathbf{e}_k = \left[ \tilde{\rho_k},\tilde\theta_k \right]^T=\left[ \sqrt{\tilde{x}_k^2+\tilde{y}_k^2},\tilde{\theta_k} \right]^T% = \\
   % \begin{bmatrix}
   %   \sqrt{(x_k^{[d]} - x_k - d\cos\theta_k^{'})^2+
   %   (y_k^{[d]} - y_k - d\sin\theta_k^{'})^2},
   %   \theta_k^{'} - \theta_k
   % \end{bmatrix},
 \end{align}
% 
 where $\theta_k^{'} = \mathrm{atan2}\left(y_k^{[d]}-y_k, x_k^{[d]}-x_k\right)$
 and $\tilde{\rho}_k =\sqrt{\tilde{x}_k^2+\tilde{y}_k^2} $ is the Euclidean distance (position error) at time instant $k$ as shown in Fig.~\ref{fig:leaderFollowerSetup} with $\tilde{x}_k = x_k^{[d]} - x_k - d\cos\theta_k^{'},$ $\tilde{y}_k = y_k^{[d]} - y_k - d\sin\theta_k^{'},$ and $\tilde{\theta}_k = \theta_k^{'} - \theta_k$ is the orientation error.  The control problem can then be formally stated as follows: Find $\nu_k$ and $\gamma_k$ such that ${\bf e}_k\to {\bf 0}$ as  $k\to\infty$ subject to~\eqref{eq:robotModelDT}~and~\eqref{eq:leaderDT}.

@@

         n = 2 , m = 2 \tcc*[h]{state error and input dimensions}\;

@@

 This section adopts the theoretical results discussed in the previous section by simulating Algorithm~\ref{alg:ModelFreeTracking} using a differential drive mobile robot. The proposed actor-critic RL algorithm is tested  utilizing MATLAB software in a laptop computer running Ubuntu 18.04 as a preliminary step to implement the Algorithm experimentally in real world. Here we present the performance of the proposed method and the convergence characteristics of the actor and critic weights. The weighting matrices are set to $\mathbf{Q} = \mathrm{diag}[0.001,0.001]$ and $\mathbf{R} = \mathrm{diag}[10^{-5}, 10^{-5}].$ % 
  % \[Q=0.001 
  % \begin{bmatrix}
  % 1       & 0   \\
  % 0       & 1   \\
  % \end{bmatrix},
  % R=0.00001
  % \begin{bmatrix}
  % 1       & 0 \\
  % 0       & 1 
  % \end{bmatrix}.\]

@@

 In the first scenario, the differential drive robot was initially placed at position $(x^d,y^d) = (-2,-2)~[\si{\meter}] $ with an orientation of $\theta = 0\degree.$ The target which is modeled as a point robot was initially located at a random position around  $(x,y) = (0,0)~[\si{\meter}].$ The simulation was run for the duration of $t_f=60~[\second].$ The target is set to move following a sinusoidal path modeled as  $
 x^{[d]}(t) = \alpha t, %
 y^{[d]}(t) = 5 \,\sin(\alpha t),$  and %
 $\theta^{[d]}(t) = cos(\alpha t),
 $ with $\alpha = 4\pi/t_f.$ and $t=kT$, $T$ represents the sampling time while $k$ being timing index. Simulation results are shown in Fig.~\ref{fig:performanceSin}. In the beginning of the simulation, the robot starts to move away from the target according to the initial random actor weights as we see in \ref{fig:trajectorySin}.In Fig \ref{fig:euclideanDistanceSin} the euclidean distance overshoots before finally approaching zero. As the robot collects more data, the robot adjust the control actions as the actor and critic weights started to converge and the robot asymptotically follows the target's trajectory as can be seen in Fig \ref{fig:criticWeightSin} and \ref{fig:actorWeightSin}. %
%
 \begin{figure*}[htbp]%

 \subfigure[][]{%
    \label{fig:trajectorySin}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/sinePath/OUT/trajectorySineWave.eps} }
	\subfigure[][]{%    
    \label{fig:euclideanDistanceSin}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/sinePath/OUT/euclideanDistanceSineWave.eps} }
%  \\
%\subfigure[][]{%
%    \label{fig:stateErrorSin}%
%    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/stateErrorRandomPathDistance.eps} }
% \subfigure[][]{%
%    \label{fig:controlInputsSin}%
%    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/secondScenario/controlInputRandomPathDistance.eps} }
  \\
 \subfigure[][]{%
    \label{fig:criticWeightSin}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/sinePath/OUT/weightSineWave.eps} }
     \subfigure[][]{%
    \label{fig:actorWeightSin}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/sinePath/OUT/weightActorSineWave.eps} }
\caption[Robot tracking a target that is  following a random trajectory.]{Robot tracking a target that is  following a random trajectory:
  \subref{fig:trajectoryRandom} robot and target trajectories,~\subref{fig:euclideanDistanceRandom} position error $\tilde\rho_k$, \subref{fig:criticWeightRandom} critic learning weights~$\bm{\omega}_c,$ and \subref{fig:actorWeightRandom} actor learning weights~${\bf W}_a$}
    \label{fig:performanceSin}%
\end{figure*} 
%

In the second scenario, the target is set to move on a completely random trajectory starting from a random initial position $(x^{[d]},y^{[d]})\approx (2,-1)~[\si{\meter}],$ where the target is now modeled as a point robot as described in ~\eqref{eq:leaderDT}
 %
 where $\mathbf{u}_k^{[d]}=[\nu^{[d]},\omega^{[d]}] $ with $\nu^{[d]}(t)=0.2~[\si{\meter}]$ and $\omega$ randomly chaning in the range of $({-\pi},{\pi}) $.  The robot initial position and orientation are $(x,y) = (0,0)~[\si{\meter}]$ and $\theta_0=0$ .We intensify that the robot has no previous knowledge of the dynamic model of the moving target. The results obtained by the simulation are shown in Fig.~\ref{fig:performanceRandom}.Figure \ref{fig:trajectoryRandom} describes the trajectory of the robot and the target during the simulation. The control actions of the robot are determined by the random actor weights. We clearly see euclidean distance in Fig \ref{fig:euclideanDistanceRandom} converges and approaches zero as actor and critic weights converges to values that allows target tracking as can be seen in fig \ref{fig:criticWeightRandom} and \ref{fig:actorWeightRandom}. The results reveal the ability of the proposed algorithm to successfully track the dynamic moving target and adapt to the changes in the target's trajectory and learns online by trial and error in different scenarios. %
%    
 \begin{figure*}[htbp]%
 \subfigure[][]{%
    \label{fig:trajectoryRandom}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/randomPath/OUT/trajectoryRandom.eps} }
	\subfigure[][]{%    
    \label{fig:euclideanDistanceRandom}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/randomPath/OUT/euclideanDistanceRandom.eps} }
%\subfigure[][]{%
%    \label{fig:stateErrorRandom}%
%    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/stateErrorRandomPathDistance.eps} }
% \subfigure[][]{%
%    \label{fig:controlInputsRandom}%
%    \includegraphics[width=0.48\textwidth,height=0.25\textheight]{figs/coppelia/firstScenario/controlInputRandomPathDistance.eps} }
 \subfigure[][]{%
    \label{fig:criticWeightRandom}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/randomPath/OUT/weightRandom.eps} }
     \subfigure[][]{%
    \label{fig:actorWeightRandom}%
    \includegraphics[width=0.48\textwidth,height=0.15\textheight]{matlabsim/randomPath/OUT/weightActorRandom.eps} }
  \caption[Robot tracking a target that is  following a random trajectory.]{Robot tracking a target that is  following a random trajectory:
  \subref{fig:trajectoryRandom} robot and target trajectories,~\subref{fig:euclideanDistanceRandom} position error $\tilde\rho_k$, \subref{fig:criticWeightRandom} critic learning weights~$\bm{\omega}_c,$ and \subref{fig:actorWeightRandom} actor learning weights~${\bf W}_a$}%
  \label{fig:performanceRandom}%
\end{figure*}
